ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-02-27 pyked.tex
added PyKED ref, and moved long URLs to footnotes

%====================================================================
\section{PyKED architecture}\label{sec:pyked-architecture}
%
\pk{} is a Python package that currently provides the reference implementation
of the interface to \ck{} files. \pk{} reads \ck{} files, validates their
structure and content, and allows the user to interact with the data contained
in the \ck{} file.

@@

\pk{} relies on well-established scientific Python software tools. These include
NumPy~\autocite{vanderWalt:2011np} and Pandas~\autocite{pandas,McKinney2010} for
array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
files, Cerberus for validating \ck{} files~\autocite{cerberus}, and
pytest~\autocite{pytest:3.0.1} for unit testing. Travis-CI~\autocite{Travis2016}
provides continuous integration testing.

\pk{} is licensed under the permissive, open-source BSD 3-clause license. The
source code is publicly available on GitHub at
\url{https://github.com/pr-omethe-us/PyKED}. Installation of stable versions can
be accomplished from a package repository, using either \pybox{pip} to install
from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
\pybox{conda} to install from the
\href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
addition, \pk{} can be installed from the source code by running
\pybox|python setup.py install| or \pybox{pip install .} from the root
directory of the repository.

@@

The \ck{} file that will be used in this example can be found in the
\pybox{tests} directory of the PyKED repository,
\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}.
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: (1, 0)
chemked-version: 0.0.1
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the data point is from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, there is just a single datapoint in this file, which codes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure, compression time, ignition type, and the volume history that specifies
how the volume of the reactor varies with time, simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
  - temperature: 297.4 kelvin
    ignition-delay: 1.0 ms
    pressure: 958.0 torr
    composition:
      - species-name: H2
        InChI: 1S/H2/h1H
        mole-fraction: 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        mole-fraction: 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        mole-fraction: 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        mole-fraction: 0.63125
    ignition-type:
      target: pressure
      type: d/dt max
    compression-time: 38.0 ms
    volume-history:
      time:
        units: s
        column: 0
      volume:
        units: cm3
        column: 1
      values:
        - [0.00E+000, 5.47669375000E+002]
        - [1.00E-003, 5.46608789894E+002]
        - [2.00E-003, 5.43427034574E+002]
        - [3.00E-003, 5.38124109043E+002]
        ...
\end{yamlbox}
%
The values for the \yabox{volume-history} are truncated to save space. A user
might like to load the information from this file to perform a simulation to
calculate the ignition delay using Cantera. All of the information required to
perform this simulation is present in the \ck{} file, with the exception of a
chemical kinetic mechanism for H$_2$/CO combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

@@

The \ck{} file used in this example can be found online.\todo{add url}{} The
data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
In this \ck{} file, multiple data points are specified with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox{common-properties} section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    - species-name: t-butanol
      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
      mole-fraction: 0.003333333
    - species-name: O2
      InChI:  1S/O2/c1-2
      mole-fraction: 0.04
    - species-name: Ar
      InChI:  1S/Ar
      mole-fraction: 0.956666667
  ignition-type:  &ign
    target: OH* concentration
    type: 1/2 max
\end{yamlbox}
%
This block utilizes the ability of YAML files to define an anchor using the
\yabox{&} symbol and refer to that section later using the \yabox{*} symbol, as
can be seen in the definition of the \yabox{datapoints} section:
%
\begin{yamlbox}
datapoints:
  - temperature: 1459 kelvin
    ignition-delay: 347 us
    pressure: 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1389 kelvin
    ignition-delay: 756 us
    pressure: 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1497 kelvin
    ignition-delay: 212 us
    pressure: 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1562 kelvin
    ignition-delay: 105 us
    pressure: 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}

>>>>

%====================================================================
\section{PyKED architecture}\label{sec:pyked-architecture}
%
\pk{} is a Python package that currently provides the reference implementation
of the interface to \ck{} files~\autocite{PyKED}. \pk{} reads \ck{} files, validates
their structure and content, and allows the user to interact with the data contained
in the \ck{} file.

@@

\pk{}~\autocite{PyKED} relies on well-established scientific Python software tools.
These include NumPy~\autocite{vanderWalt:2011np} and Pandas~\autocite{pandas,McKinney2010}
for array manipulation, Pint~\autocite{Grecco2016} for interpreting and converting
between units, \texttt{ruamel.yaml}~\autocite{ruamel_yaml} for parsing YAML
files, Cerberus for validating \ck{} files~\autocite{cerberus}, and
pytest~\autocite{pytest:3.0.1} for unit testing. Travis-CI~\autocite{Travis2016}
provides continuous integration testing.

\pk{} is licensed under the permissive, open-source BSD 3-clause license. The
source code is publicly available on GitHub at
\url{https://github.com/pr-omethe-us/PyKED}~\autocite{PyKED}. Installation of stable versions
can be accomplished from a package repository, using either \pybox{pip} to install
from \href{https://pypi.python.org/pypi}{PyPI} (Python Package Index), or
\pybox{conda} to install from the
\href{https://anaconda.org/conda-forge}{\texttt{conda-forge}} channel. In
addition, \pk{} can be installed from the source code by running
\pybox|python setup.py install| or \pybox{pip install .} from the root
directory of the repository.

@@

The \ck{} file that will be used in this example can be found in the
\pybox{tests} directory of the PyKED
repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_rcm.yaml}}.
Examining that file, we find the first section specifies the information about
the \ck{} file itself:
%
\begin{yamlbox}
file-author:
  name: Kyle E Niemeyer
  ORCID: 0000-0003-4425-7097
file-version: (1, 0)
chemked-version: 0.0.1
\end{yamlbox}
%
Then, we find the information regarding the article in the literature from which
this data was taken. In this case, the data point is from the work of
\textcite{Mittal2006a}:
%
\begin{yamlbox}
reference:
  doi: 10.1002/kin.20180
  authors:
    - name: Gaurav Mittal
    - name: Chih-Jen Sung
      ORCID: 0000-0003-2046-8076
    - name: Richard A Yetter
  journal: International Journal of Chemical Kinetics
  year: 2006
  volume: 38
  pages: 516-529
  detail: Fig. 6, open circle
experiment-type: ignition delay
apparatus:
  kind: rapid compression machine
  institution: Case Western Reserve University
  facility: CWRU RCM
\end{yamlbox}
%
Finally, there is just a single datapoint in this file, which codes the ignition
delay for the experiment, the mixture composition, the initial temperature,
pressure, compression time, ignition type, and the volume history that specifies
how the volume of the reactor varies with time, simulating the compression
stroke and post-compression processes.
%
\begin{yamlbox}
datapoints:
  - temperature: 297.4 kelvin
    ignition-delay: 1.0 ms
    pressure: 958.0 torr
    composition:
      - species-name: H2
        InChI: 1S/H2/h1H
        mole-fraction: 0.12500
      - species-name: O2
        InChI: 1S/O2/c1-2
        mole-fraction: 0.06250
      - species-name: N2
        InChI: 1S/N2/c1-2
        mole-fraction: 0.18125
      - species-name: Ar
        InChI: 1S/Ar
        mole-fraction: 0.63125
    ignition-type:
      target: pressure
      type: d/dt max
    compression-time: 38.0 ms
    volume-history:
      time:
        units: s
        column: 0
      volume:
        units: cm3
        column: 1
      values:
        - [0.00E+000, 5.47669375000E+002]
        - [1.00E-003, 5.46608789894E+002]
        - [2.00E-003, 5.43427034574E+002]
        - [3.00E-003, 5.38124109043E+002]
        ...
\end{yamlbox}
%
The values for the \yabox{volume-history} are truncated to save space. A user
might like to load the information from this file to perform a simulation to
calculate the ignition delay using Cantera. All of the information required to
perform this simulation is present in the \ck{} file, with the exception of a
chemical kinetic mechanism for H$_2$/CO combustion.
%
\begin{pythonbox}
import cantera as ct
from pyked import ChemKED

@@

The \ck{} file used in this example can be found in the \pybox{tests} directory of the \pk{} 
repository\footnote{\url{https://github.com/pr-omethe-us/PyKED/blob/master/pyked/tests/testfile_st_p5.yaml}}.
The data in this file comes from Stranic et al.~\autocite{Stranic:2012}, describing
shock-tube ignition delays for \textit{tert}-butanol. We have omitted the file
meta information below for space; the format is largely similar to the example in Section~\ref{sec:rcm-modeling}.
%
In this \ck{} file, multiple data points are specified with some common
conditions, including a common mixture composition and common definition of
ignition delay. Therefore, a \yabox{common-properties} section is specified.
%
\begin{yamlbox}
common-properties:
  composition: &comp
    - species-name: t-butanol
      InChI: 1S/C4H10O/c1-4(2,3)5/h5H,1-3H3
      mole-fraction: 0.003333333
    - species-name: O2
      InChI:  1S/O2/c1-2
      mole-fraction: 0.04
    - species-name: Ar
      InChI:  1S/Ar
      mole-fraction: 0.956666667
  ignition-type:  &ign
    target: OH* concentration
    type: 1/2 max
\end{yamlbox}
%
This block utilizes the ability of YAML files to define an anchor using the
\yabox{&} symbol and refer to that section later using the \yabox{*} symbol, as
can be seen in the definition of the \yabox{datapoints} section:
%
\begin{yamlbox}
datapoints:
  - temperature: 1459 kelvin
    ignition-delay: 347 us
    pressure: 1.60 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1389 kelvin
    ignition-delay: 756 us
    pressure: 1.67 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1497 kelvin
    ignition-delay: 212 us
    pressure: 1.55 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
  - temperature: 1562 kelvin
    ignition-delay: 105 us
    pressure: 1.50 atm
    composition: *comp
    ignition-type: *ign
    equivalence-ratio: 0.5
\end{yamlbox}
